comgen-outline

### 2017-12-07 paper/comgen-paper.Rmd
edits to other philosophical puzzles section

The uniform prior over the threshold $\theta$ in the truth-functional threshold semantics (as we have assumed in our model) provides a secondary argument for why acquiring our semantics for generics should be easy.
Uniform uncertainty over values in the unit interval $[0, 1]$ for threshold $\theta$ in a threshold-function is mathematically equivalent to a *soft semantics* wherein the degree to which the utterance is true is proportional to the feature-probability $h$ *itself*: 
$$
\int_{0}^{1} \delta_{h > \theta} \diff \theta =  h
$$

This *soft semantics* (corresponding intuitively to a meaning like "the higher the probability, the better the utterance", or simply "more is better") is perhaps the simplest semantics one could imagine.
<!-- Given this mathematical equivalence, we can imagine the *more is better* soft semantics is what is acquired early. -->
The difficulty in acquiring the meaning of quantifiers is then a difficulty in recognizing a fixed-threshold semantics as a special case of this *more is better* semantics.
We leave for future work the precise implementation of the acquisition model we sketch out here.
<!-- This special case involves learning the value of the threshold (aspect 2) and not in recognizing that the utterance conveys something about the quantity (e.g., aspect 1). -->
## Other philosophical puzzles

In our experiments, we have tried to include major known philosophically puzzling examples of generalizations in language. 
In this section, we consider other previously-discussed troubling examples of generic statements and sketch out arguments rooted in the theoretical framework presented in this paper.

@@

(1) is often used to demonstrate that the feature-probability is not sufficient for a generic to be true.
The vast majority of books that exist are paperbacks; nonetheless *Books are paperbacks* sounds weird.
This is not an isolated example.
The same features can be observed in *Pastas are spaghetti*, *Animals are humans*.
These examples are generated by pairing a superordinate category with a property that describes a strictly subordinate category. 
We posit that the weirdness of these examples is the result of a failure of the comparison class.
<!-- Specifically, the referent-category K in a generic statement "Ks F" is not expected to be in the comparison class.  -->

Though we have not introduced a theory of how a comparison class is contructed, relevant ideas may be found in @Keil1979.
His notion of *predicability* might relate to the comparison class, insofar as both may be derived with respect the property (e.g., the kinds of categories can be felicitiously predicated by "is paperback" may form a comparison class).
Note that this idea is similar to @Cohen1999's category of "Relative generics", which have truth conditions that are determined by a comparison to an alternative set of Ks (his Alt(K), our comparison class) which is composed of Ks that could satisfy some F in an an alternative set of Fs.
For example, being a paperback can only be true or false (in the predicability sense of @Keil1979) of individual books.
Aggregations of entities that can be predicated by the feature are, without further supporting context, categories of books.
Such categories may include books of different of different genres or eras, among plausibly many other such sub-categories.
From the taxonomic relation of the subordinate categories (e.g., types of books) to the referent-category (i.e., books), the true probability of the feature among the referent-category must be the mean of the probabilities of all subcategories, assuming an equal partitioning of the prior probability of subcategories.
Thus, if we were to construct a speaker model of the statement "Books are paperbacks", the model would predict a 0.5 endorsement: neither true nor untrue, in the same way as "Robins are female".

(2) is often used to argue that no actual instances need have the property to make a generalization true. 
In this thought experiment, part (perhaps all) of Mary's job is to handle the mail from Antarctica; to date, however, there has not been any mail from Antarctica [@Cohen1999]. 
Thus, it is not the case that Mary *has handled* any mail from Antarctica before.
This example is an extreme case of the phenomenon highlighted by Expts. 2c \& 2d.
In this pair of experiments, we manipulated predictive probability, while keeping fixed past frequency, and saw the endorsements of habitual statements indeed tracked predictive probability (with the linking function being the RSA speaker model). 
Our general cognitive modeling framework takes the fundamental currency of a generalization in language to be a subjective predictive probability. 
Past frequency and actual prevalence in the world often tracks predictive probabilities, but language users' internal models of how the world works can lead them in certain situations to expect something different.

The focus on subjective, predictive probability casts new light on classic puzzles surrounding accidentally-true situations.
Consider a thought experiment from the linguistics and philosophical literature: Imagine that currently every Supreme Court Justice's social security number is an even number. 
Linguists predict the statement "Supreme Court Justices have even social security numbers" would still be considered false [@Cohen1999].
Our explanation of this phenomenon is that abstract intuitive theories lead us to reject observed frequencies in forming our subjective probabilities.
That is, because we may believe selection for the Supreme Court is not influenced by one's social security number, we would assign a roughly 50\% subjective probability that the \emph{next} justice would have an even social security number.
Thus, given that the all professions would have roughly the same probability of having employees with even social security numbers, the statement \emph{Supreme Court Justices have even social security numbers} would be similar to *Birds are female*. 
Thus, our model would not endorse the statement \emph{Supreme Court Justices have even social security numbers}, because the predictive feature-probability would not be any different for Supreme Court Justices than any other profession.
This perspective makes the intriguing prediction that if we learned much more surprising information, we might be compelled to revise our theory of the domain, update our subjective probability of future instances, and then accept the generic. 
For instance, if every justice in history had \emph{prime numbered} social security numbers (a more suspicious coincidence), one might appeal to a conspiracy, which \emph{would} have predictive consequences.

(3) shows a subtle interaction of generic language with another linguistic device: conjunction. ^[
This example was first introduced, to our knowledge, by @Nickel2008.
]
@Nickel2008 suggests this statement is semantically equivalent to: *Elephants live in Africa and elephants live in Asia*.
This example is challenging for many statistical based approaches, since it cannot be the case that most elephants live in Africa and most elephants live in Asia. In addition, we should not take the statement to imply that there are international elephants (i.e., individual elephants who live in Africa part time and Asia part time); but that is the only possible interpretation if the generic statement is given a fixed, majority based quantificational account (e.g., if the generic means *most*).

@@

With reasonable assumptions about the prior (e.g., that *lives in* is similar to other biological properties from Expt. 1a and assuming international elephants are unlikely *a priori*), upon the hearing "Elephants live in Africa", our model will infer that most elephants live in Africa.
When the model then hears "Elephants live in Asia", the model updates its beliefs to accomodate this second statement, inferring that some live in Africa and other lives in Asia. 
This is possible because our model has uncertainty over the semantic thresholds in the generic claim.
Our model can flexibly accomodate new evidence (e.g., learning that elephants live in Asia) that might otherwise contradict previous statements (e.g., knowing that elephants live in Africa).
A runnable version of the model for this example can be found at [http://forestdb.org/models/generics-conjunction.html](http://forestdb.org/models/generics-conjunction.html).

Finally, (4) is used to argue that even if the property is widespread (or perhaps in a softer sense, as in our model, widespread relative to other kinds or groups), speakers might find it avervise to actually assert the generic.
That is, even if it were the case that, because of a variety of socio-cultural reasons, women tend to be submissive (more so than men, and many other social categories), we might still not want to assert *Women are submissive* [@Haslanger2011]. 
In our Model Simulations section, we showed that the generic often can imply the property is widespread, even with uninformed prior beliefs about the property. 
If the generic conveys the property is in fact more widespread than the speaker herself might believe, this can lead to a distortion of the truth. ^[
Note that this can naturally capture the central phenoman described in @Cimpian2010.
].
The current formulation of the model is relatively ambivalent about the deviations of the listener's beliefs from the speaker's belief (i.e., the speaker doesn't mind that the feature-probability that she believes to be true will be exaggerated from the generic statement).
However, if states of the world (or, listener's beliefs about states of the world) correspond to different subjective utilities, speakers who take into the account the subjective utilities of the listener may produce language that deviates from what a pure informational transmission speaker might say.
This kind of extension to the RSA framework has already accounted for a variety of behaviors corresponding to polite speech understanding and production [@Yoon2016; @Yoon2017].
If our speaker model believed that states of the world where the probability of the feature is high (e.g., a world where many women were submissive) were undesirable (either in absolute terms or as a belief), then a speaker who considered both the informational content of the utterance and the utility in the states of the world implied by that utterance would too find it reprehensible to make such generic claims. 
We find this possibility intruiging for future work.

In sum, we've tried to argue that our information-theoretic communicative model couched in general cognitive framework can be used to formalize a number of interesting and heretofore puzzling avenues in genericity. 
The precise implementation of these hypotheses we leave for future work.

>>>>

The uniform prior over the threshold $\theta$ in the truth-functional threshold semantics (as we have assumed in our model) provides a secondary argument for why acquiring our semantics for generics should be easy.
Uniform uncertainty over values in the unit interval $[0, 1]$ for threshold $\theta$ in a threshold-function is mathematically equivalent to a *soft semantics* wherein the degree to which the utterance is true is proportional to the feature-probability $h$ *itself*:

$$
\int_{0}^{1} \delta_{h > \theta} \diff \theta =  h
$$
This *soft semantics* (corresponding intuitively to a meaning like "the higher the probability, the better the utterance", or simply "more is better") is perhaps the simplest semantics one could imagine.
The difficulty in acquiring the meaning of quantifiers is then a difficulty in recognizing a fixed-threshold semantics as a special case of this *more is better* semantics.
We leave for future work the precise implementation of the acquisition model we sketch out here.
## Other philosophical puzzles

In our experiments, we have tried to include classic, philosophically puzzling examples of generalizations in language in addition to crafting novel stimuli that provide stricter tests of the main hypothesis.
In this section, we consider some other previously-discussed troubling examples of generalizations and sketch out arguments rooted in our theoretical framework to demonstrate how our theory could account for these cases. 
We consider the five following examples:

@@

Sentence (1), strikes hearers as *odd* or perhaps *false*, despite the vast majority of books being paperback. 
Why?
A first observation is that the sentence is constructed by pairing a category ("Books") with a property that describes a strict subset ("paperbacks" is short-hand for "paperback books"). 
We expect intuitions about this statement to generalize to other examples expressing an analogous relation (e.g., "Pastas are spaghetti", "Animals are humans").
We hypothesize that the strangeness of these examples has to do an inflexible relationship between the comparison class and the referent feature-probabilty.
Assume that the comparison class is constructed by considering the *predicability* of the property [in the sense of @Keil1979, originally from @Sommers].^[
  Note that this assumption is consistent with the comparison classes we have assumed in our three case studies.
]
Then, "is a paperback" can only be informatively said of individual books, because these are the only objects that can plausibly have the property; that is, world knowledge gives us that an elephant cannot be predicated (either true or falsely) by the property "is a paperback".^[
  Note that this idea is similar to @Cohen1999's notion of "Relative generics", which have truth conditions that are determined by a comparison to an alternative set of Ks (his   Alt(K), our comparison class) which is composed of Ks that could satisfy some F in an an alternative set of Fs.
]
The relevant aggregations of entities, then, are collection of books, potentially including books of different genres or publication eras among other such sub-categories.
Then, given that "books" is the superordinate category of "kinds of books", the referent-category feature-probability cannot be different from the mean of the feature-probability prior.
Finally, given this feature-probability prior and the referent-category feature-probability, our endorsement model would predict "Books are paperbacks" is neither true nor untrue, in the a way analogous to "Robins are female".

```{r booksArePaperbacks, fig.cap="Sketch of an argument for why 'Books are paperbacks' sounds odd. The feature 'is paperback' can only be predicated (in the sense of @Keil1979) of individual books, which are organized into collections (here, by genre). The feature-probability prior is constructed by aggregating these collections of books. Then the feature-probability of 'is paperbook' for the category of 'books' can never be higher than the mean of the prior distribution. The endorsement model would give this sentence a rating of 0.5, corresponding to neither true nor false."}
knitr::include_graphics("figs/books-paperbacks.pdf", dpi = 108)
```

Imagine there is a job in the local beaurocrat's office to handle the mail from Antarctica and this job is assigned to Mary;  to date, however, there has never been any mail from Antarctica [@Cohen1999]. 
That is, the statement "Mary has handled mail from Antarctica" is false.
Sentence (2) is still thought to be true despite zero actual instances of the event.
We think (2) is a boundary case of generalizations expressing predictive probabilities, which we elucidated in Expt. 2c.
In in that study, we experimentally dissociated predictive probability from past frequency, and saw that only a model that took into account predictive probability could explain participants' endorsements of generalizations.
We expect the same to be true here, elucidated by the intuition that: were there to be any mail coming from Antarctica, Mary would handle it. 
Past frequency or actual prevalence in the world often tracks predictive probabilities, but language users' internal models of how the world works can lead them in certain situations to expect something different in the future.

Our focus on predictive probability makes even new prediction about some classic puzzles.
Sentence (3) is the conceptual opposite of sentence (2).
Imagine that currently every Supreme Court Justice has a social security number which is an *even number*. 
Sentence (3) is still considered false, every though the property holds in 100\% of the category [@Cohen1999].
We predict this phenomenon is the result of abstract, intuitive theories guiding us to reject observed frequencies in forming our subjective probabilities.
That is, because observers strongly believe selection for the Supreme Court is not influenced by one's social security number, speakers would assign a roughly 50\% subjective probability that the \emph{next} justice would have an even social security number.
Then, given that the all professions would have roughly the same probability of having employees with even social security numbers, Sentence (3) is again similar to "Birds are female". 
However, were we to learn much more surprising information---for instance, if every Supreme Court Justice in U.S. history had a social security number which was a *prime number*---the shear suspiciousness could compel an observer to revise their theory of the domain (appealing perhaps to a conspiracy), update their subjective probability of future instances, and then accept the generic. 
That is, we predict there are situations where statements like (3) could be considered felicitious generalizations.

Understanding how a semantic representation *composes* is another critical test for a theory of generalizations in language.
@Nickel2008 suggests Sentence (4) is semantically equivalent to: "Elephants live in Africa and elephants live in Asia".
Stastistical account of generalizations have trouble with this kind of conjunction: It cannot be the case that *most* (more than half) elephants live in Africa and *most* (more than half) elephants live in Asia, unless we are to posit *international elephants* (i.e., individual elephants who live part-time in Africa and part-time in Asia), which we would rather not have to posit.
However, this is the only interpretation if the generalization is given a fixed, majority based quantificational account (e.g., if the generic means *most*).
Our theory does not provide a fixed semantics for the generalizations, but an uncertain one, which can be updated as more information comes in.
In fact, with a strong constraint against the existence of *international elephants* but with any further semantic assumptions, our model infers that some elephants live in Africa and other lives in Asia.^[
  An implementation of this example can be found: [http://forestdb.org/models/generics-conjunction.html](http://forestdb.org/models/generics-conjunction.html).
]
Importantly, the model believes that most lives in Africa, when it only hears that "Elephants live in Africa", and only later revises it's beliefs to a weaker "some"-like interpretation after hearing they live in Asia as well.
Our model flexibly accomodates new evidence that might otherwise contradict previous utterances because it maintains uncertainty about the precise interpretation of single utterances.

@@

<!-- With reasonable assumptions about the prior (e.g., that *lives in* is similar to other biological properties from Expt. 1a and assuming international elephants are unlikely *a priori*), upon the hearing "Elephants live in Africa", our model will infer that most elephants live in Africa. -->
<!-- When the model then hears "Elephants live in Asia", the model updates its beliefs to accomodate this second statement,  -->
<!-- This is possible because our model has uncertainty over the semantic thresholds in the generic claim. -->

Finally, sometimes a property is widespread relative to other kinds or groups, and it might be so systemic that we would believe it to be the case in the future, but speakers still might find it avertise to actually endorse the generalization.
Such is the situation with Sentence (5) [@Haslanger2011]. 
Due of a variety of socio-cultural reasons, @Haslanger2011 argues that women tend to be submissive (more so than men, and many other social categories), yet still cautions against asserting (5).
In our **Model Simulations** section, we showed that the genearlization often implies the property is widespread, even with relatively uninformed prior beliefs about the property. 
If the generic conveys the property is in fact more widespread than the speaker herself might believe, this can lead to a distortion of the truth. ^[
  Note that this can straight-forwardly capture the central phenomena of @Cimpian2010.
].
The current formulation of the model is relatively ambivalent about the deviations of the listener's beliefs from the speaker's belief (i.e., the speaker doesn't mind that the feature-probability that she believes to be true will be exaggerated in the mind of the listener).
However, if states of the world (or, listener's beliefs about states of the world) correspond to different subjective values, speakers who take into the account the subjective utilities of the listener may produce language that deviates from what a pure informational transmission speaker might say.
This kind of extension to rational models of communication has already been used to account for a variety of phenomena in *polite language* use [@Yoon2016; @Yoon2017].
We propose adopting such a value-based utility structure for states of the world for language about social categories: If the speaker found states of the world where the probability of the feature is high (e.g., a world where many women were submissive) as undesirable, then a speaker who considered both the informational content of the utterance and the utility in the states of the world implied by that utterance would too find it reprehensible to make such a generalization.
We find this possibilitity intruiging and leave it for future work.

In sum, we've tried to argue that our information-theoretic communicative model couched in general cognitive framework can be used to explain a number of interesting and heretofore puzzling phenomena in genericity as well as make novel predictions about a wider range of language.
The precise implementation of these hypotheses we leave for future work.
